Cosmology with photometric surveys: constraints with redshift tomography of non--Gaussian statistics

### 2016-03-30 photoz_paper/paper.tex
completed first draft

\begin{abstract}
Weak Gravitational Lensing is becoming a popular technique to constrain cosmological parameters, such as the Dark Energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for a LSST like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We show that different statistics lead to different bias directions in parameter space, leaving a window open for self--calibration. 
\end{abstract}

@@

\section{Introduction}
%
This work is organized as follows: we give an outline of the shear simulations we use in this work, followed by descriptions of the convergence reconstruction procedure and forward modeling of galaxy shape and redshift systematics. We then give an overview of the parameter inferences technique we used to forecast constraints on cosmology. We then discuss the main results of this work and present our conclusions as well as prospects for future work.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxy have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}. The parameter space sampling we adopted for our simulations is shown in Figure \ref{fig:simdesign}

\begin{figure*}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{Parameter space sampling}
\label{fig:simdesign}
\end{figure*}
%
For each of the parameter choices in Figure \ref{fig:simdesign}, the $N$--body initial conditions are generated using the same random seed. In addition to this simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number of pseudo--independent catalog realizations is produced.

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/galdistr.eps}
\caption{Galaxy distribution assumed throughout this work, along with the choice of the redshift bins $\bar{z}_b$.}
\label{fig:galdistr}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scibook.eps}
\caption{LSST Science Book}
\label{fig:lsstscibook}
\end{figure}

@@

\subsection{Convergence reconstruction}
In this paragraph we describe the procedure we used to construct convergence maps $\kappa$ from the simulated shear catalogs $\pmb{\gamma}$. We consider a two dimensional square pixel grid of area $\theta_{\rm FOV}^2$ and with 512 pixel per side. This correspond to a linear pixel resolution of $0.5 \rm{arcmin}$. We assign a shear value $\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b)$ to each pixel $\pmb{\theta}_p$ according to the following procedure
\begin{equation}
\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b) = \frac{\sum_{g=1}^{N_g}\pmb{\gamma}(\pmb{\theta}_g,z_g)W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b)}{\sum_{g=1}^{N_g}W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b)}
\end{equation}   
%
We chose a top--hat window function
\begin{equation}
W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b) = 
\begin{cases}
1 \,\,\,\,{\rm if}\,\,\,\, \pmb{\theta}_g\in\pmb{\theta}_p,z_g\in\bar{z}_b \\
0 \,\,\,\,{\rm otherwise}
\end{cases}
\end{equation} 
%
The convergence $\kappa(\pmb{\theta}_p)$ can be reconstructed from the $E$--mode of the shear field, which is evaluated from the Fourier Transform of the pixelized shear $\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b)$
\begin{equation}
\label{meth:psdefinition}
\tilde\kappa(\pmb{\ell},\bar{z}_b) = \left(\frac{\tilde{\gamma}^1(\pmb{\ell},\bar{z}_b)(\ell_x^2-\ell_y^2)+2\ell_1\ell_2\tilde{\gamma}^2(\pmb{\ell},\bar{z}_b)}{\ell_x^2+\ell_y^2}\right) e^{-\frac{\ell^2\theta_G^2}{2}}
\end{equation}
%
We chose the Gaussian filter smoothing scale $\theta_G=0.5\,{\rm arcmin}$ to correspond to the linear pixel resolution. Inverting the Fourier Transform yields the pixelized map $\kappa(\pmb{\theta}_p,\bar{z}_b)$. We apply this procedure to both the shear realizations $\h{\pmb{\gamma}}_r(\pmb{\theta}_g,z_g;\bb{p})$ and the mock observations $\h{\pmb{\gamma}}_{\rm obs}(\pmb{\theta}_g,z_g)$, yielding convergence realizations $\h{\kappa}_r(\pmb{\theta}_p,\bar{z}_b;\bb{p})$ and mock convergence observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$.

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the $\kappa$ tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
{\rm Re}\left(\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle\right) = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropical, the expectation value $\langle\rangle$, for each realization $r$, is taken over all the modes $\pmb{\ell}$ which have the same magnitude $\ell=\vert\pmb{\ell}\vert$. The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the $\kappa$ local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the $\kappa$ standard deviation over pixels. The $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$ are defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}) 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c)
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity, and also because both peak counts and moments do not contain cross redshift information unlike the power spectrum defined in equation (\ref{meth:psdefinition}). In the next paragraph we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from mock observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$, $\pmb{\mu}(\bar{z}_b)$.

@@

\subsection{Dimensionality reduction}
We apply Principal Component Analysis techniques (see \citep{astroMLText} for example) to reduce the dimensionality of our summary statistics while preserving the cosmological information content. Consider the model statistic $\bb{d}(\bb{p})$, and think of it in terms of a $P\times N_b$ matrix $d_{pi}$. Consider the whitened model matrix
\begin{equation}
\Delta_{pi} = P\frac{d_{pi}}{\sum_{q=1}^Pd_{qi}} - 1
\end{equation} 
%
We perform a Singular Value Decomposition (SVD) of $\Delta$ 
\begin{equation}
\bb\Delta = \bb{L}\bb{\Lambda} \bb{R}
\end{equation}
%
where $L$ is $P\times Q$, $\bb{\Lambda}={\rm diag}(\Lambda_1,...,\Lambda_Q)$ and $\bb{R}$ is $Q\times N_b$ and $Q={\min}(P,N_b)$. $R_{ni}$ is the $i$-th component of the $n$-th basis vector in statistics space. The singular value $\Lambda_n$ is the variance of the whitened summary statistic along the $n$-th basis vector. We assume that only summary statistic projections on the first $N_c$ basis vectors contain relevant cosmological information, and $N_c<N_b$ is a number that has to be determined from the simulations. Let $\bb{R}(N_c)$ be a matrix made of the first $N_c$ rows of $\bb{R}$ (we assume that the singular values $\Lambda_i$ are sorted). We define the PCA projection of a summary statistic $\bbh{d}$ on $N_c$ principal components as 
\begin{equation}
\h{d}_n(N_c) = \sum_{i=1}^{N_b}(\bb{R}(N_c))_{ni}\left(P\frac{\h{d}_i}{\sum_{p=1}^P d_{pi}}-1\right) 
\end{equation} 
%
In this way we hope to capture the cosmological information contained in $\bbh{d}$ by projecting it on the $N_c<N_b$ principal components that vary the most with cosmology parameters. In the next section we outline the main results of this work.  
%%%%%%%%%%%%%%%%%%%%%%% RESULTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}

\begin{figure*}
\includegraphics[scale=0.3]{Figures/w_power_spectrum_pca.eps}
\includegraphics[scale=0.3]{Figures/w_peaks_pca.eps}
\includegraphics[scale=0.3]{Figures/w_moments_pca.eps}
\caption{Constraints on $w$ as a function of $N_c$}
\label{fig:pcacomponents}
\end{figure*}

\begin{figure}
\includegraphics[scale=0.3]{Figures/photoz_bias.eps}
\caption{Bias due to photo-$z$ errors}
\label{fig:photozbias}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/constraints.eps}
\caption{Constraints in $(\Omega_m,w)$ space}
\label{fig:constraintsOm-w}
\end{figure}

>>>>

\begin{abstract}
Weak Gravitational Lensing is becoming a popular technique to constrain cosmological parameters, such as the Dark Energy equation of state $w$. When analyzing galaxy surveys, redshift information has proven to be a valuable addition to angular shear correlations. We forecast parameter constraints on the triplet $(\Omega_m,w,\sigma_8)$ for a LSST like galaxy photometric survey, using tomography of the shear-shear power spectrum, convergence peak counts and higher convergence moments. We find that the addition of peak and counts to the power spectrum can significantly improve the constraints on the $(\Omega_m,w)$ doublet. We evaluate the effect of uncorrected photometric redshift systematics on parameter constraints. We find that different statistics lead to different bias directions in parameter space, leaving a window open for self--calibration. 
\end{abstract}

@@

\section{Introduction}
%
Weak Gravitational Lensing is a promising technique to probe the Large Scale Structure of the universe in which the tracers are unbiased \citep{wlreview}. This technique has the potential of improving significantly the constraints on the Dark Energy equation of state $w$ because it is most sensitive to the matter density fluctuations at the non--linear stage. Cosmology inferences from Weak Lensing observations have been produced for past (CFHTLenS \citep{cfht1}) and current (DES \citep{DES}) surveys, and are being planned for future experiments as well (LSST \citep{LSST}, WFIRST \citep{WFIRST}, Euclid \citep{Euclid}). Because of the non-linear nature of the density fluctuations probed by Weak Lensing, cosmological information might leak from quadratic statistics (such as two point functions and power spectra) into more complicated non-Gaussian statistics, for which forward modeling from cosmological parameters requires numerical simulations of cosmic shear fields. An extensive list of these non-Gaussian statistics, and their cosmological information content, have been investigated extensively in the literature. We investigate the constraining power of a subset of these non-Gaussian statistics combined with redshift tomography in a LSST-like survey. We also investigate the effects of uncorrected photometric redshift systematics on parameter constraints when using redshift tomography. This work is organized as follows: we give an outline of the shear simulations we use in this work, followed by descriptions of the convergence reconstruction procedure and forward modeling of galaxy shape and photometric redshift systematics. We then give an overview of the parameter inferences techniques we used to forecast constraints on cosmology. We then discuss the main results of this work and present our conclusions as well as prospects for future work.

@@

\subsection{Cosmic shear simulations}
\label{sec:shearsim}
We review the procedure used for generating mock shear catalogs. We consider a fiducial flat $\Lambda$CDM universe with $(h,\Omega_m,\Omega_\Lambda,\Omega_b,w,\sigma_8,n_s)=(0.72,0.26,0.74,0.046,-1,0.8,0.96)$. We examine different variations of the $\bb{p}=(\Omega_m,w,\sigma_8)$ triplet and run one $N$--body simulation for each choice of $\bb{p}$, using the public code \ttt{Gadget2} \citep{Gadget2}. The simulations have a comoving box size of $L_b=260\,{\rm Mpc}/h$ and contain $512^3$ dark matter particles, which correspond to a mass resolution of $M_p\approx 10^{10}M_{\rm sun}$ per particle. The three dimensional outputs of the $N$--body simulations are sliced in sequences of two dimensional lenses $120 \,{\rm Mpc}$ thick, which are lined up perpendicular to the line of sight between the observer on Earth and a source at redshift $z_s$. We make use of the multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} to trace the deflections of light rays originating at $z=0$ through the system of lenses until $z$. To accomplish this task, we make use of the \ttt{LensTools} \citep{LensTools-ASCL,LensTools-paper} implementation of the multi--lens--plane algorithm. An observed galaxy position $\pmb{\theta}$ on the sky today corresponds to a real galaxy angular position $\pmb{\beta}(\pmb{\theta},z_s)$, which can be calculated using the \ttt{LensTools} pipeline by solving the lensing ODE up to redshift $z_s$. The Jacobian of $\pmb{\beta}(\pmb{\theta},z_s)$ is a $2\times 2$ matrix that contains information about the cosmic shear field at $\pmb{\theta}$ integrated along the line of sight. 
%
\begin{equation}
\label{meth:sheareqn}
\frac{\partial\beta_i(\pmb{\theta},z_s)}{\partial \theta_j} = 
\begin{pmatrix}
1-\kappa(\pmb{\theta})-\gamma^1(\pmb{\theta}) & -\gamma^2(\pmb{\theta}) \\
-\gamma^2(\pmb{\theta}) & 1-\kappa(\pmb{\theta})+\gamma^1(\pmb{\theta})\\
\end{pmatrix}
\end{equation}  
%
The quantities that appear in equation (\ref{meth:sheareqn}) are the convergence $\kappa$, which is the source magnification due to lensing, and the cosmic shear $\pmb{\gamma}$, which is a measurement of the source ellipticity due to lensing from Large Scale Structure, assuming the non lensed shape is a circle. 
We simulate $N_g = 10^6$ random galaxy positions $\{\pmb{\theta}_g\}$ distributed uniformly in a field of view of size $\theta_{\rm FOV}^2=(3.5{\rm deg})^2$, which correspond to a galaxy angular density of $n_g=22\,\rm{arcmin}^{-2}$. The galaxy have a distribution in redshift which mimics the one expected in the LSST survey
\begin{equation}
\label{meth:galdistr}
n(z) = n_0\left(\frac{z}{z_0}\right)^2\exp\left(-\frac{z}{z_0}\right)
\end{equation}  
%
with $z_0=0.3$ and $n_0$ a normalization constant fixed so that $n(z)$ integrates to the total number of galaxies $N_g$. For each galaxy we compute the cosmic shear at $\pmb{\theta}_g$ using equation (\ref{meth:sheareqn}), producing a shear catalog $\{\pmb{\gamma}_g\}$. Different random realizations of a shear catalog $\{\pmb{\gamma}_g\}_r$ can be obtained rotating and periodically shifting the Large Scale Structure in the $N$--body snapshots according to the procedure explained in \citep{PetriVariance}. We produce $N_r=16000$ pseudo--independent realizations of the shear catalog $\{\pmb{\gamma}_g\}$. These shear realization cover the total survey area of LSST 10 times.
We repeat the above procedure for $P=100$ different combinations of the parameter triplet $\bb{p}$, sampled according to a Latin hypercube scheme. The sampling procedure is the same described in \citep{CFHTMink,CFHTPeaks}. The parameter space sampling we adopted for our simulations is shown in Figure \ref{fig:simdesign}

\begin{figure*}
\includegraphics[scale=0.4]{Figures/design.eps}
\caption{Parameter space sampling chosen for our simulations. We show both the sampling in the $(\Omega_m,w)$ (left) and $(\Omega_m,\sigma_8)$ (right) slices of the parameter space. The fiducial cosmology has been marked in red.}
\label{fig:simdesign}
\end{figure*}
%
For each of the parameter choices in Figure \ref{fig:simdesign}, the $N$--body initial conditions are generated using the same random seed. In addition to this simulations, we produce simulated shear catalogs for a fiducial $\Lambda$CDM universe with $\bb{p}_0=(0.26,-1,0.8)$. In this case the randomization procedure is based on 5 independent $N$--body simulations, and the same number of pseudo--independent catalog realizations is produced.

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/galdistr.eps}
\caption{Galaxy distribution assumed throughout this work (see equation (\ref{meth:galdistr})), along with the choice of the redshift bins $\bar{z}_b$. Our galaxy sample consists of $N_g=10^6$ galaxies. The figure shows the number of galaxies $N_g(z)$ at each redshift $z$.}
\label{fig:galdistr}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scibook.eps}
\caption{Photo-$z$ bias $b_{\rm ph}(z_s)$ (black) and standard deviation $\sigma_{\rm ph}(z_s)$ (red) as a function of the source redshift $z_s$ for two different models, optimistic (solid lines) and pessimistic (dashed lines), obtained from the LSST Science Book \citep{LSSTSciBook}.}
\label{fig:lsstscibook}
\end{figure}

@@

\subsection{Convergence reconstruction}
In this paragraph we describe the procedure we used to construct convergence maps $\kappa$ from the simulated shear catalogs $\pmb{\gamma}$. We consider a two dimensional square pixel grid of area $\theta_{\rm FOV}^2$ and with 512 pixel per side. This correspond to a linear pixel resolution of $0.5 \rm{arcmin}$. We assign a shear value $\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b)$ to each pixel $\pmb{\theta}_p$ according to the following procedure
\begin{equation}
\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b) = \frac{\sum_{g=1}^{N_g}\pmb{\gamma}(\pmb{\theta}_g,z_g)W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b)}{\sum_{g=1}^{N_g}W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b)}
\end{equation}   
%
We chose a top--hat window function
\begin{equation}
W(\pmb{\theta}_g,\pmb{\theta}_p;z_g,\bar{z}_b) = 
\begin{cases}
1 \,\,\,\,{\rm if}\,\,\,\, \pmb{\theta}_g\in\pmb{\theta}_p,z_g\in\bar{z}_b \\
0 \,\,\,\,{\rm otherwise}
\end{cases}
\end{equation} 
%
The convergence $\kappa(\pmb{\theta}_p)$ can be reconstructed from the $E$--mode of the shear field, which is evaluated from the Fourier Transform of the pixelized shear $\pmb{\gamma}(\pmb{\theta}_p,\bar{z}_b)$
\begin{equation}
\label{meth:psdefinition}
\tilde\kappa(\pmb{\ell},\bar{z}_b) = \left(\frac{\tilde{\gamma}^1(\pmb{\ell},\bar{z}_b)(\ell_x^2-\ell_y^2)+2\ell_x\ell_y\tilde{\gamma}^2(\pmb{\ell},\bar{z}_b)}{\ell_x^2+\ell_y^2}\right) e^{-\frac{\ell^2\theta_G^2}{2}}
\end{equation}
%
We chose the Gaussian filter smoothing scale $\theta_G=0.5\,{\rm arcmin}$ to correspond to the linear pixel resolution. Inverting the Fourier Transform yields the pixelized map $\kappa(\pmb{\theta}_p,\bar{z}_b)$. We apply this procedure to both the shear realizations $\h{\pmb{\gamma}}_r(\pmb{\theta}_g,z_g;\bb{p})$ and the mock observations $\h{\pmb{\gamma}}_{\rm obs}(\pmb{\theta}_g,z_g)$, yielding convergence realizations $\h{\kappa}_r(\pmb{\theta}_p,\bar{z}_b;\bb{p})$ and mock convergence observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$.

We measure a variety of summary statistics from the pixelized convergence maps, which will then be used to forecast parameter constraints and biases. We consider three kinds of summary statistics, namely the $\kappa$ tomographic power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, the tomographic peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and a set of $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$. The tomographic power spectrum is defined as 
\begin{equation}
\langle\tilde{\kappa}(\pmb{\ell},\bar{z}_b)\tilde{\kappa}(\pmb{\ell}',\bar{z}_{b'})\rangle = (2\pi)^2\delta_D(\pmb{\ell}+\pmb{\ell}')P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})
\end{equation}
%
Because the $\kappa$ field is statistically isotropical, the expectation value $\langle\rangle$, for each realization $r$, is taken over all the modes $\pmb{\ell}$ which have the same magnitude $\ell=\vert\pmb{\ell}\vert$. The peak count statistic $n_{\rm pk}(\nu,\bar{z}_b)$ is defined as the number of the $\kappa$ local maxima of a certain height $\kappa_{\rm max}=\nu\sigma_0$, where $\sigma_0$ is the $\kappa$ standard deviation over pixels. The $\kappa$ moments $\pmb{\mu}(\bar{z}_b)$ are defined as follows (see \citep{Matsubara10,Munshi12,MinkPetri}) 
%
\begin{equation}
\label{meth:momdef}
\begin{matrix}
\pmb{\mu} = (\pmb{\mu}_2,\pmb{\mu}_3,\pmb{\mu}_4) \\ \\
\pmb{\mu}_2 = (\langle\kappa^2\rangle,\langle\vert\nabla\kappa\vert^2\rangle) \\ \\
\pmb{\mu}_3 = (\langle\kappa^3\rangle,\langle\kappa\vert\nabla\kappa\vert^2\rangle,\langle\kappa^2\nabla^2\kappa\rangle) \\ \\
\pmb{\mu}_4 = (\langle\kappa^4\rangle_c,\langle\kappa^2\vert\nabla\kappa\vert^2\rangle_c,\langle\kappa^3\nabla^2\kappa\rangle_c,\langle\vert\nabla\kappa\vert^4\rangle_c)
\end{matrix}
\end{equation}
%
In equation (\ref{meth:momdef}) the gradients $\nabla$ are evaluated using finite differences between $\kappa$ values at neighboring pixels and the expectation values $\langle\rangle$ for each realization $r$ are taken over the $512^2$ pixels in the map. The subscript $c$ indicates that we consider only the connected parts of the quartic $\kappa$ moments. In the definition of the peak counts and convergence moments we omitted the redshift index $\bar{z}_b$ for notational simplicity, and also because both peak counts and moments do not contain cross redshift information unlike the power spectrum defined in equation (\ref{meth:psdefinition}). In the next paragraph we describe the statistical methods we use to infer cosmological parameter estimates $\bbh{p}$ from mock observations $\h{\kappa}_{\rm obs}(\pmb{\theta}_p,\bar{z}_b)$ using the summary statistics $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, $n_{\rm pk}(\nu,\bar{z}_b)$, $\pmb{\mu}(\bar{z}_b)$.

@@

\subsection{Dimensionality reduction}
We apply Principal Component Analysis techniques (see \citep{astroMLText} for example) to reduce the dimensionality of our summary statistics while preserving the cosmological information content. Consider the model statistic $\bb{d}(\bb{p})$, and think of it in terms of a $P\times N_b$ matrix $d_{pi}$. Consider the whitened model matrix
\begin{equation}
\Delta_{pi} = P\frac{d_{pi}}{\sum_{q=1}^Pd_{qi}} - 1
\end{equation} 
%
We perform a Singular Value Decomposition (SVD) of $\Delta$ 
\begin{equation}
\bb\Delta = \bb{L}\bb{\Lambda} \bb{R}
\end{equation}
%
where $L$ is $P\times Q$, $\bb{\Lambda}={\rm diag}(\Lambda_1,...,\Lambda_Q)$ and $\bb{R}$ is $Q\times N_b$ and $Q={\min}(P,N_b)$. $R_{ni}$ is the $i$-th component of the $n$-th basis vector in statistics space. The singular value $\Lambda_n$ is the variance of the whitened summary statistic along the $n$-th basis vector. We assume that only summary statistic projections on the first $N_c$ basis vectors contain relevant cosmological information, and $N_c<N_b$ is a number that has to be determined from the simulations. Let $\bb{R}(N_c)$ be a matrix made of the first $N_c$ rows of $\bb{R}$ (we assume that the singular values $\Lambda_i$ are sorted). We define the PCA projection of a summary statistic $\bbh{d}$ on $N_c$ principal components as 
\begin{equation}
\label{meth:pcaprojection}
\h{d}_n(N_c) = \sum_{i=1}^{N_b}(\bb{R}(N_c))_{ni}\left(P\frac{\h{d}_i}{\sum_{p=1}^P d_{pi}}-1\right) 
\end{equation} 
%
In this way we hope to capture the cosmological information contained in $\bbh{d}$ by projecting it on the $N_c<N_b$ principal components that vary the most with cosmology parameters. In the next section we outline the main results of this work.  
%%%%%%%%%%%%%%%%%%%%%%% RESULTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results}
In this section we illustrate the main results of this work. Figure \ref{fig:pcacomponents} shows the robustness of the dimensionality reduction technique we adopted. on the three summary statistics considered, namely the convergence power spectrum $P^{\kappa\kappa}(\ell,\bar{z}_b,\bar{z}_{b'})$, peak counts $n_{\rm pk}(\nu,\bar{z}_b)$ and moments $\pmb{\mu}(\bar{z}_b)$. To measure the power spectrum we chose 15 uniformly spaced multipole bins between $(\ell_{\rm min},\ell_{\rm max})=(200,2000)$, for a total of $N_b=15\times 15=225$ measurement bands, including cross redshift information. We bin the convergence peak counts in 30 uniformly spaced $\nu$ bins between $(\nu_{\rm min},\nu_{\rm max})=(-2,7)$, for a total of $N_b=30\times5=150$ measurement bands, including redshift information. The total size of the moments summary statistic vector is $N_b=9\times 5=45$, including redshift information. The forecast error bars on $w$ are calculated according to equation (\ref{meth:parcovestimator}), where the covariance matrix $\bbh{C}$ and its inverse $\bbh{\Psi}$ have been estimated with $N_r=16000$ realizations of each summary statistic in the fiducial cosmology. Figure \ref{fig:constraintsOm-w} shows the $1\sigma$ confidence contours on the $(\Omega_m,w)$ doublet calculated with equation (\ref{meth:parcovestimator}) after the PCA dimensionality reduction performed according to equation (\ref{meth:pcaprojection}), for a variety of choices $({\rm statistic},N_c)$. We also show the improvements on the confidence contours when combining different summary statistics after the corresponding dimensionality reductions have been performed. Figure \ref{fig:photozbias} shows the effect of ignoring photo-$z$ errors on parameter constraints. To evaluate this effect we construct different mock observations, with and without photo-$z$ effects, and compare the results of the parameter fit according to equation (\ref{meth:parestimate}). Using our simulation suite, we construct 20 mock observations: the summary statistic in each observation is calculated by taking the mean of a random sample of $N_{\rm FOV}$ realizations of the summary statistic in the fiducial cosmology (randomly chosen among the ensemble of $N_r=16000$ that are available in the ensemble). The estimated covariance matrix $\bbh{C}$ is scaled by a factor $N_{\rm FOV}$ to take into account the construction process of the mock observations. This procedure allows to forecast the results an LSST-like survey would obtain.

\begin{figure*}
\includegraphics[scale=0.3]{Figures/w_power_spectrum_pca.eps}
\includegraphics[scale=0.3]{Figures/w_peaks_pca.eps}
\includegraphics[scale=0.3]{Figures/w_moments_pca.eps}
\caption{$1\sigma$ errors on $w$ as a function of the number of the principal components $N_c$, using the power spectrum (top left), peak counts (top right) and moments (bottom). The thin colored lines refer to single redshift summary statistics, while the thick black line shows the case in which the joint redshift information is included.}
\label{fig:pcacomponents}
\end{figure*}

\begin{figure}
\includegraphics[scale=0.3]{Figures/constraints.eps}
\caption{$1\sigma$ confidence contours on the $(\Omega_m,w)$ parameter space obtained using equation (\ref{meth:parcovestimator}). The covariance matrix $\bbh{C}$ and its inverse $\bbh{\Psi}$ have been obtained using $16000$ summary statistics realizations, and has been scaled by a factor $N_{\rm FOV}=1600$ to mimic the constraining power of an LSST-like survey.}
\label{fig:constraintsOm-w}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/photoz_bias.eps}
\caption{Distribution of parameter estimates in the $(\Omega_m,w)$ parameter space using the power spectrum (black), peak counts (black) and moments (blue) to fit 20 LSST-like mock observations. The parameter deviations $(\delta\Omega_m,\delta w)$ are obtained comparing parameter estimates according to equation (\ref{meth:parestimate}) with and without photo-$z$ errors. Results for both optimistic (dots) and pessimistic (crosses) are shown. The colored squares and the ellipses correspond to respectively to the mean and $1\sigma$ level of the $(\delta\Omega_m,\delta w)$ distribution in the pessimistic systematics case, assuming this distribution is Gaussian.}
\label{fig:photozbias}
\end{figure}
